---
title: "Project2-Exploratory"
author: "group2"
date: "2021/7/25"
output: pdf_document
---

```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA)
library(dplyr) ; library(moderndive); library(tidyverse) ; library(GGally); library(skimr); library(kableExtra); library(ggplot2);library(gridExtra)
```
```{r dataset}
household <- read_csv("dataset2.csv")
household <- household%>%select(-Region)
```
# Exploratory Data Analysis {#sec:EDA}

## Summary Analysis
```{r Summary}
my_skim <- skim_with(base = sfl(n = length))
household%>%
  select(-Household.Head.Sex, -Type.of.Household, -Electricity)%>%
  my_skim()%>%
  transmute(Variable=c('Total Household Income', 'Total Household Expenditure', 'Household Head Age', 'Number of Family Members', 'Floor Area', 'House Age', 'Number of Bedrooms'), n=n, Mean=numeric.mean, SD=numeric.sd, Min=numeric.p0, Q1=numeric.p25, Median=numeric.p50,Q3=numeric.p75, Max=numeric.p100)%>%
  kable(caption = '\\label{tab:summary} Summary statistics', digits=2)%>%
  kable_styling(latex_options = "hold_position")
```

According to the descriptive statistic, there is not any missing value in this database and the number of observations is 1249. There are large standard deviations observed in total household income and total household expenditure, which means the range between rich and poor in this survey is quite considerable. The avarage capacity of a family is more than 4 people, which ranges from 1 to 16. Most houses have a floor area lies in 5 to 60 square meters, while some houses are drastically big and can be measured up to 750 square meters. Three quarters of the houses are under 22 years old. However, the oldest house could be dated back to 105 years ago. The number of bedrooms ranges from 0 to 7, which has an average of 1.78.

## Barplots
```{r boxplot1, out.width = '68%', fig.align = "center", fig.cap = "\\label{fig:box} Number of Family Members by Family Head Sex.", fig.pos = 'H'}
#boxplot of family head sex
ggplot(household, aes(x = Household.Head.Sex, y = Total.Number.of.Family.members)) +
geom_boxplot() +
labs(x = "Household Head Sex", y = "Total Number of Family Members")
```
The households that have a male head tend to have a larger family, i.e. the median and interquatile range of male head household are larger compared to female head household.

```{r boxplot2, out.width = '68%', fig.align = "center", fig.cap = "\\label{fig:box} Number of Family Members by Type of Household.", fig.pos = 'H'}
#boxplot of Type of Household
ggplot(household, aes(x = Type.of.Household, y = Total.Number.of.Family.members)) +
geom_boxplot() +
labs(x = "Type of Household", y = "Total Number of Family Members")
```
The extended families accomodate more family members compared to other two types, which have a larger median as well. Single families are slightly smaller but still larger than families made with two or more nonrelated persons.

```{r boxplot3, out.width = '68%', fig.align = "center", fig.cap = "\\label{fig:box} Number of Family Members by Electricity.", fig.pos = 'H'}
#boxplot of Electricity
ggplot(household, aes(group=Electricity, x = Electricity, y = Total.Number.of.Family.members)) +
geom_boxplot() +
labs(x = "Electricity", y = "Total Number of Family Members")
```
The influence of electricity to family capacity is not clear since the median values are same between households with and without electricity. However, the interquatile range of no electricity families is larger which means more minor families live without electricity.
```{r boxplot4, out.width = '68%', fig.align = "center", fig.cap = "\\label{fig:box} Number of Family Members by Number of Bedrooms.", fig.pos = 'H'}
#boxplot of bedrooms
ggplot(household, aes(group=Number.of.bedrooms, x = Number.of.bedrooms, y = Total.Number.of.Family.members)) +
geom_boxplot() +
labs(x = "Number of Bedrooms", y = "Total Number of Family Members")
```
The median family members in household with five bedrooms shows the largest, which is 5, while which stay 4 in households have under five bedrooms. Households with five bedrooms also have the largest interquatile range, shows a big difference in family capacities. Nevertheless, households with six bedrooms shows the lowest median and range.

## Scatter Plots
```{r ScatterPlots}
p1 <- ggplot(household, aes(y = Total.Number.of.Family.members,
                            x = Total.Household.Income)) +
  geom_point(alpha=0.3) +
  labs(y = "Family Members", x = "Total Household Income") +
  geom_smooth(method = "lm", se = F)
p2 <- ggplot(household, aes(y =Total.Number.of.Family.members ,
                            x =Total.Food.Expenditure)) +
  geom_point(alpha=0.3) +
  labs(y = "Family Members", x = "Food Expenditure") +
  geom_smooth(method = "lm", se = F)
p3 <- ggplot(household, aes(y = Total.Number.of.Family.members,
                            x = Household.Head.Age)) +
  geom_point(alpha=0.3) +
  labs(y = "Family Members", x = "Head Age") +
  geom_smooth(method = "lm", se = F)
p4 <- ggplot(household, aes(y = Total.Number.of.Family.members,
                            x = House.Floor.Area)) +
  geom_point(alpha=0.3) +
  labs(y = "Family Members", x = "House Floor Area") +
  geom_smooth(method = "lm", se = F)
p5 <- ggplot(household, aes(y = Total.Number.of.Family.members,
                            x = House.Age	)) +
  geom_point(alpha=0.3) +
  labs(y = "Family Members", x = "House Age") +
  geom_smooth(method = "lm", se = F)
p6 <- ggplot(household, aes(y =  Total.Number.of.Family.members,
                            x =Number.of.bedrooms	)) +
  geom_point(alpha=0.3) +
  labs(y = "Family Members", x = "Number of Bedrooms") +
  geom_smooth(method = "lm", se = F)
grid.arrange(p1,p2,p3,p4,p5,p6, nrow = 2)
```
(I don't know whether this plot need into use so I didn't interpret it)

## ggplots
```{r ggpairs,out.height='100%', out.width= "100%", fig.align='center', fig.cap="\\label{fig:ggpairs1} Correlation Plot ", fig.pos= "H"}
ggpairs(household[,c(-3,-5,-10)],axisLabels="none", columnLabels=c('Income','Food','Head Age', 'Members','Floor Area','House Age', 'Bedrooms'),lower=list(continuous=wrap("points", alpha=0.4)))
```

The correlation between household food expenditure and household income is large, which indicates the multicollinearity may need take into consideration. (bedroom&income/food/floor area?) According to the correlation in this graph, the number of family members has a moderate positive relationship with household food expenditure, while other relationships are not obvious and need further research.

# Formal Analysis

Let's have a look at the GLM Poisson model of Response against all Explanatory variables.

```{r echo=FALSE,message=FALSE}
library(moderndive) ; library(tidyverse) ; library(GGally) ; library(skimr) ; library(plotly)
library(jtools) ; library(kableExtra) ; library(ggfortify) ; library(janitor) ; library(infer)
library(gridExtra) ; library(olsrr) ; library(data.table) ; library(knitr) ; library(corrplot)
library(ggcorrplot) ; library(gapminder) ; library(sjPlot) ; library(stats) ; library(jtools)
library(MASS) ; library(smacof) ; library(broom) ; library(ROCR) ; library(nnet)
library(arm) ; library(ggstance) ; library(effects) ; library(COMPoissonReg) ; library(Rcpp)
library(AER)
```


```{r Model Poisson}
dataset2 <- read_csv("dataset2.csv")
str(dataset2)
dataset2 <- dataset2 %>% dplyr::select(-Region)
dataset2$Electricity <- as.factor(ifelse(dataset2$Electricity == "1", "yes", "No"))
dataset2$Household.Head.Sex <- as.factor(dataset2$Household.Head.Sex)
mod_poi_full <- glm(Total.Number.of.Family.members ~ Total.Household.Income + Total.Food.Expenditure +
                 Household.Head.Age + House.Floor.Area + House.Age + Number.of.bedrooms +
                 Type.of.Household + Household.Head.Sex + Electricity,
                family = poisson, data = dataset2)
summary(mod_poi_full)

```

Here it can been that few explanatory variables have insignificant p value. And confident intervals also contains zero so Let's work on these variables. Let's look at individual models, response against explanatory variables.

```{r}
mod_sample2 <- glm(Total.Number.of.Family.members ~ Total.Household.Income,
                  family = poisson, data = dataset2)
#summary(mod_sample2)

mod_sample3 <- glm(Total.Number.of.Family.members ~ Total.Food.Expenditure,
                  family = poisson, data = dataset2)
#summary(mod_sample3)

mod_sample4 <- glm(Total.Number.of.Family.members ~ Household.Head.Age,
                  family = poisson, data = dataset2)
#summary(mod_sample4)

mod_sample5 <- glm(Total.Number.of.Family.members ~ Type.of.Household,
                  family = poisson, data = dataset2)
#summary(mod_sample5)

mod_sample6 <- glm(Total.Number.of.Family.members ~ Household.Head.Sex,
                  family = poisson, data = dataset2)
#Summary(mod_sample6)

mod_poi <- glm(Total.Number.of.Family.members ~ Total.Household.Income +
                 Household.Head.Age + Number.of.bedrooms + Household.Head.Sex,
                family = poisson,data = dataset2)

summary(mod_poi)
```

From this analysis we can see that the veriables given defines the model better so we will go with these variables and let's have a look at model fit.
```{r}
mod_poi$deviance
Mean <- mean(dataset2$Total.Number.of.Family.members)
Variance <- var(dataset2$Total.Number.of.Family.members)
c(Var = Variance, Mean = Mean) %>%
  kable(col.names = c("Response"), digits = 3, caption = '\\label{tab:Stats} Mean and Variance of Response variable') %>%
  kable_styling(latex_options = "striped")
```

```{r Variance Comparison}
ggplot(mod_poi, aes(x=log(fitted(mod_poi)), y=log((dataset2$Total.Number.of.Family.members-fitted(mod_poi))^2)))+
  geom_point(col="#f46d43") +
  geom_abline(slope=1, intercept=0, col="#a6d96a", size=1) +
  ylab(expression((y-hat(mu))^2)) + xlab(expression(hat(mu)))
```

The residual deviance of model is 1247.3 and degree of freedom is 1244. Mean and Variance of response variable have a little difference of 0.4 which can indicate to overdispersion. Furthermore the plot between fitted values and $\epsilon^2$ we can see that there is no problem of dispersion as the residuals have almost equal spread agaist the line. We will proceed to with chi square test, and dispersion test from *AER* library to see how good is our model.

## Goodness of fit

```{r chi square test}
D <- mod_poi$null.deviance - mod_poi$deviance
chi_sq <- qchisq(p = 0.95, df = mod_poi$df.null-mod_poi$df.residual)
c("D0-D1" = D, "X^2 (10)" = chi_sq) %>%
  kable(col.names = c("values"), digits = 3, caption = '\\label{tab:Chi_sq test 1} Chi Square Test: Null Model vs Complete Model') %>%
  kable_styling(latex_options = "striped")

chi_sq2 <- qchisq(p = 0.95, df = mod_poi$df.residual)
c("resid D" = mod_poi$deviance, "X^2 (1238)" = chi_sq2) %>%
  kable(col.names = c("values"), digits = 3, caption = '\\label{tab:Chi_sq test 2} Chi Square Test: Residual Deviance') %>%
  kable_styling(latex_options = "striped", )


disp_test <- dispersiontest(mod_poi, trafo = NULL, alternative = c("greater", "two.sided", "less"))

X2 <- sum(resid(mod_poi, type = "pearson")^2)
disp_par <- X2 / mod_poi$df.residual

c(disp_test$estimate, "Disp Parameter" = disp_par) %>%
  kable(col.names = c("values"), digits = 3, caption = '\\label{tab:Dispersion} Dispersion') %>%
  kable_styling(latex_options = "striped", )
```

From the comparison of null model and full model in table \ref{tab:Chi_sq test 1} we can see that the $\chi^2(4)$ value is less than the difference between null deviance and residual deviance. which can indicate lack of fit. Secondly from the table \ref{tab:Chi_sq test 2} we can see that residual deviance is greater than $\chi^2(1244)$, which can also lead to lack of fit. lastly, we can see that dispersion test and dispersion parameter value is same So we can try Quasipoisson using **dispersion parameter**.

## Dispersion Parameter

$$\hat{\phi} = \frac{X^2}{n - p}$$

As because of dispersion parameter Wald statistic is not valid, so to check the significance of coefficients we go for "**F**" test.

```{r}
summary(mod_poi, dispersion =disp_par )

drop1(mod_poi, test = "F")
confint(mod_poi)
```

According to the confidence intervals and F statistic, we can see that there is no more insignificant variables. To deal with a little bit of dispersion in model we go for Quasipoisson of negative binomial model and observe results

## Quasipoisson and Negative Binomial model

```{r}
mod_qpoi <- glm(Total.Number.of.Family.members ~ Total.Household.Income +
                 Household.Head.Age + Number.of.bedrooms + Household.Head.Sex,
                family = quasipoisson, data = dataset2)

summary(mod_qpoi)
mod_nb <- glm.nb(Total.Number.of.Family.members ~ Total.Household.Income +
                 Household.Head.Age + Number.of.bedrooms + Household.Head.Sex,
                data = dataset2)

summary(mod_nb)
```

We can compare the Poisson and Negative Binomial models by looking at their deviances and AIC scores:

```{r model_comparisons}
c(mod_qpoi$deviance, mod_qpoi$aic)
c(mod_nb$deviance, mod_nb$aic)
```

Both are much smaller for the negative binomial model so the Negative Binomial model is preferred over the Poisson model.

```{r}
results <- cbind(Estimate = coef(mod_nb), confint(mod_nb))
exp(results) %>% kable(digits = 3) %>% kable_styling(latex_options = "striped")
```
